Repeatability of Pre-Flashover Fire Patterns on Gypsum Wall Board

### 2015-03-08 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Add figures to Madrzykowski thesis.

Examples of the photographic images are presented in Figs.~\ref{Gas_Foam_Photos} and \ref{Gasoline_Photos}.  Figure 14 is a photograph of the natural gas flame taken during a period of ``steady state'' burning at approximately 80~kW and a photograph of the polyurethane foam fire taken near its peak heat release rate of approximately 60~kW.  Figure 9 presents two photographs of the same gasoline fueled fire when it was burning near its peak heat release rate of 100~kW.  The two photographs, taken 0.2~s apart, represent the range of the observed flame heights from one of the gasoline experiments from 0.65~m and 1.0~m, respectively.

@@

The calculations for the intensity field are based on a user defined computation grid.  A grid sensitivity analysis was conducted with 5~mm, 10~mm, 20~mm, 40~mm and 80~mm square grids.  Examples of grid cell resolution outputs from Streams for the natural gas fueled burner are shown in Figure 19.  The outputs shown are time averaged intensity fields generated with 5 resolutions listed above. In each output the x and y axis dimensions are in mm.

@@

Another means of quantifying field data with Streams is the use of contour plots.  Examples of the time averaged contour plots are shown in Figures 20 through 22 for a natural gas, gasoline, and polyurethane fueled experiments respectively.  Each contour line represents the percentage of time that the flame was at a given height.  Figure 20 provides an example of a natural gas experiment.  The region under examination, as shown as a black bounding rectangle, is 1400~mm high and 800 mm in width.  This example was averaged over approximately 29~s or approximately 870 images.  The number above each contour line in the plot range from 0 to 1.  The area bounded by the contour with the value 1 represents the area that the flame occupied 100~\% of the time period analyzed or the persistent flame region as defined by McCaffrey~\cite{McCaffrey:1979}. The remaining contours are in the intermittent flame region, with the 0.5 contour representing the mean flame height. Outside of the final contour, value 0.0, no visible flame was recorded; hence this contour marks the definitive start of the buoyant plume region.

Figures 22 and 23 show two different contours for the polyurethane foam fueled experiment.  The results from polyurethane fueled fire characterization experiments have an increased amount of variability due to the nature of the ignition source and the fuel itself.  The polyurethane foam was readily ignited with a small pilot flame similar to that used to ignite the natural gas and the gasoline.  By comparison, a small flame ignited the gases flowing out the top of the natural gas burner and the flames from the natural gas cover entire burner surface within approximately one second.  A similar ignition sequence occurs with the gasoline pan fires. The polyurethane foam has a more complicated flame spread sequence that resulted in significant changes to the peak heat release rate and the time to peak heat release.  Once the polyurethane foam was ignited, it took approximately 60~s for the flame to spread across the top surface of the fuel.  As the flames spread horizontally across the fuel surface, the fire also burns down into the fuel.  In some cases the fire burned completely through the fuel in the center of the foam sample, resulting in a hollow flame at the time of the peak heat release rate.  An example of a contour plot, exhibiting the hollow or bifurcated flames is shown in Figure 23.  Other methods of ignition for the polyurethane foam were considered in order to have a more uniform ignition of the top surface of the foam.  Each method would have resulted in a larger release of energy from the ignition source or a starter fuel.   Based on a few scoping experiments these methods did not improve the repeatability.

>>>>

Examples of the photographic images are presented in Figs.~\ref{Gas_Foam_Photos} and \ref{Gasoline_Photos}.  Figure~\ref{Gas_Foam_Photos} is a photograph of the natural gas flame taken during a period of ``steady state'' burning at approximately 80~kW and a photograph of the polyurethane foam fire taken near its peak heat release rate of approximately 60~kW.  Figure~\ref{Gasoline_Photos} presents two photographs of the same gasoline fueled fire when it was burning near its peak heat release rate of 100~kW.  The two photographs, taken 0.2~s apart, represent the range of the observed flame heights from one of the gasoline experiments from 0.65~m and 1.0~m, respectively.

@@

The calculations for the intensity field are based on a user defined computation grid.  A grid sensitivity analysis was conducted with 5~mm, 10~mm, 20~mm, 40~mm and 80~mm square grids.  Examples of grid cell resolution outputs from Streams for the natural gas fueled burner are shown in Fig.~\ref{Intensity}.  The outputs shown are time averaged intensity fields generated with 5 resolutions listed above. In each output the x and y axis dimensions are in mm.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{../Figures/Fig19}\\
  \caption[Comparison images of intensity fields derived from the video images of the natural gas fueled fire]{Comparison images of intensity fields derived from the video images of the natural gas fueled fire using Streams. The images, starting with the upper left, were generated with 5 mm, 10 mm, 20 mm, 40 mm and 80 mm resolutions.}
  \label{Intensity}
\end{figure}

@@

Another means of quantifying field data with Streams is the use of contour plots.  Examples of the time averaged contour plots are shown in Figs.~\ref{Gas_and_Gasoline_Contours} and \ref{Foam_Contours} for the natural gas, gasoline, and polyurethane fueled fires.  Each contour line represents the percentage of time that the flame was at a given position. The region under examination, as shown as a black bounding rectangle, is 1400~mm high and 800~mm in width.  The natural gas fire was sampled for 29~s, resulting in 870~images. The area bounded by the contour with the value 1 represents the area that the flame was visible 100~\% of the time sampled. This can be considered the persistent flame region as defined by McCaffrey~\cite{McCaffrey:1979}. The remaining contours are in the intermittent flame region, with the 0.5 contour representing the mean flame height. Outside of the final contour, value 0.0, no visible flame was recorded; hence this contour marks the definitive start of the buoyant plume region.

\begin{figure}
  \includegraphics[width=2.7in]{../Figures/Fig20} 
  \includegraphics[width=2.7in]{../Figures/Fig21} \\
  \caption[Visibility fractions for the natural gas and gasoline fires]{Plots showing the fraction of time that the flame was visible for the natural gas (left) and gasoline (right) fires.}
  \label{Gas_and_Gasoline_Contours}
\end{figure}

\begin{figure}
  \includegraphics[width=2.7in]{../Figures/Fig22} 
  \includegraphics[width=2.7in]{../Figures/Fig23} \\
  \caption[Visibility fractions for the polyurethane foam fire]{(Left) Plot showing the fraction of time that the flame was visible for the polyurethane foam fire. (Right) Polyurethane foam contour plot with bifurcated flames at time of peak heat release rate.}
  \label{Foam_Contours}
\end{figure}

Figure~\ref{Foam_Contours} shows two different contours for the polyurethane foam fueled experiment.  The results from polyurethane fueled fire characterization experiments have an increased amount of variability due to the nature of the ignition source and the fuel itself.  The polyurethane foam was readily ignited with a small pilot flame similar to that used to ignite the natural gas and the gasoline.  By comparison, a small flame ignited the gases flowing out the top of the natural gas burner and the flames from the natural gas cover entire burner surface within approximately one second.  A similar ignition sequence occurs with the gasoline pan fires. The polyurethane foam has a more complicated flame spread sequence that resulted in significant changes to the peak heat release rate and the time to peak heat release.  Once the polyurethane foam was ignited, it took approximately 60~s for the flame to spread across the top surface of the fuel.  As the flames spread horizontally across the fuel surface, the fire also burns down into the fuel.  In some cases the fire burned completely through the fuel in the center of the foam sample, resulting in a hollow flame at the time of the peak heat release rate.  An example of a contour plot, exhibiting the hollow or bifurcated flames is shown at right in Fig.~\ref{Foam_Contours}.  Other methods of ignition for the polyurethane foam were considered in order to have a more uniform ignition of the top surface of the foam.  Each method would have resulted in a larger release of energy from the ignition source or a starter fuel.   Based on a few scoping experiments these methods did not improve the repeatability.
